### 2018-01-12 proposal/proposal.tex
First draft of proposal (and related files)

\documentclass{article}

\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage[hyperref=true, natbib=true, style=numeric, backend=bibtex]{biblatex}
\usepackage{graphicx}
\bibliography{references.bib}

\setlength{\parindent}{0pt}
\setlength{\parskip}{\baselineskip}

\begin{document}

\title{Practical Passphrase Cracking}
\date{\today{}}
\author{Luc Gommans \\
	Radically Open Security
}

\maketitle

\section{Introduction}

Most of user data is password protected, but the choice of these passwords is
largely left to users. Password recommendations propose using hard-to-remember
patterns in order to make it harder to guess, while in reality many of those
`random' tweaks (such as a number or exclamation mark at the end) are easily
iterable by a computer.

Using passphrases, users can approach the problem differently: form a string of
random, common words. The most famous example is `correct horse battery
staple': four words with no coherence, but many people in the field can recall
this phrase without ever trying to memorize it\cite{own}. Studies also show
that users have fewer difficulties when memorizing a passphrase than when
memorizing equally-strong random
passwords\cite{behavioral-analysis}\cite{pwd-memorability}.

Labrande\cite{crackmeimfamous} showed that at least 4 million in a set of 139
million passwords (2.9\%) consist of phrases from publicly available sources.
The exact percentage of passphrases used as passwords is not known, but this
shows they are used by a lof of users while they have been studied relatively
little.

Password cracking tools are widely available, which help us understand the
strength of them. For passphrases, there is no commonly used approach using a
single tool or dataset, probably because there are very few of them.

In this work, we will attempt to create a program which generates likely
passphrases. It can be used in combination with traditional tools such as
Hashcat or John the Ripper. We will use a probabilistic method to create
realistic sentences and use rulesets of the traditional tool to do additional
manipulation such as uppercasing the first word, removing spaces, and other
operations that can be done using the optimized code of that tool.

\section{Related Work}

Sparell and Simovits\cite{sparell-simovits} created a passphrase cracking
program by using Markov chains on a word level. Markov chains are used in many
applications, such as next word prediction in keyboards on smartphones.
Similarly, one can use them to approximate the next most likely word in a
passphrase given a few starting words (which could be a dictionary walk). In
Sparell and Simovits' work, the LinkedIn hashdump of 2012 was used as one of
the test sets. This contains mostly passwords (not phrases), but is still
frequently used due to its large size of 64 million hashes. Of these, 21 000
phrases were cracked of between 10 and 20 characters in length.

In previous work, we\cite{own} generated passphrases from quotes of various
well-known persons and song lyrics. Using this method on the LinkedIn set, we
recovered 92 000 passphrases between 15 and 67 characters. A relatively large
amount of time was spent on collecting this data, which we think could be
optimized in future work. We also looked into whether users incorporate
personal data, such as lyrics specifically from a band they enjoy, but we have
not found this to be the common case.

Labrande\cite{crackmeimfamous} attempted to crack hashes from the Korelogic
dump (139 million entries) used a compilation of different public sources.
Around 4.3 million passwords were cracked, of which 200 000 are 16 characters
or longer. Public sources include quotes from WikiQuote and text written in
bold and italics on Wikipedia.

To summarize, the problem of cracking passphrases is generally approached by
either using a large set of phrases directly, or by training a model on a large
set and using those to derive phrases which are likely to be logical.

\section{Research Questions}

Our main research question is: {\it How can software efficiently generate
likely passphrases, to be used in passphrase cracking?}

To achieve this, we will use a probabilistic method which generates phrases in
decreasing likelihood. An important factor in this design is the speed at which
it can do this: most methods are not designed to iteratively list all
possibilities in decreasing likelihood, which makes it difficult to do so using
methods such as context-free grammar (as described by
Weir\cite{probabilistic-thesis}, p. 84--85).

From this follow more questions to answer our main question:

\begin{enumerate}
	\item Which probabilistic methods can be used to generate output in decreasing order of likelihood?
	\item Which probabilistic method produces the most likely passphrases?
	\item How much data, both in terms of storage and random access memory, does this method require?
\end{enumerate}

Our questions focus on probabilistic methods because those do not require users
to download or even compile large lists of passphrases to use directly.
Instead, a model can be made available which contains probabilities for a
certain language.

\section{Scope}

The software to be created will interface with at least one common password
cracking tool, such as Hashcat or John the Ripper, and should be easy to use in
concert. It is considered out of scope to implement cracking in the software
itself, since that functionality can well be provided by other software.

The software will work for a general population and cannot be primed with facts
such as birth dates of specific users. In our previous work\cite{own}, we found
this to be not as effective as expected. It seems better to focus on the
general method in this limited-time project.

\section{Requirements}

For this project, a software prototype will be built which might use a
medium-sized dataset to operate. We estimate that a system with 4GiB RAM and
10GiB persistent storage is required. This is supplied by Radically Open
Security.

For software, we expect to use only free and open source products.

\section{Ethical Considerations}

In this investigation, we will need to evaluate the prototype. We will do this
using passwords from publicly available datasets. In all cases, the usernames
will be removed and the only information kept is the password (or a hash
thereof) and how frequently it occurred in the dataset.

\section{Planning}

\begin{tabular}{ | l | l | }
\hline {\bf Week 1} & Write proposal, choose probabilistic method \\
\hline {\bf Week 2} & Early prototype, interfacing with password cracking software \\
\hline {\bf Week 3} & Improving and evaluating prototype \\
\hline {\bf Week 4} & Finalise prototype and paper, prepare presentation \\
\hline
\end{tabular}

\printbibliography

\end{document}

### 2018-01-12 proposal/proposal.tex
Proposal: processed feedback

@@
 
-Most of user data is password protected, but the choice of these passwords is
-largely left to users. Password recommendations propose using hard-to-remember
-patterns in order to make it harder to guess, while in reality many of those
-`random' tweaks (such as a number or exclamation mark at the end) are easily
-iterable by a computer.
-
-Using passphrases, users can approach the problem differently: form a string of
+Requirements for secure passwords are getting more and more stringent. Using
+passphrases, users can approach the problem differently: form a string of
 random, common words. The most famous example is `correct horse battery
@@
 passwords\cite{behavioral-analysis}\cite{pwd-memorability}.
-
-Labrande\cite{crackmeimfamous} showed that at least 4 million in a set of 139
-million passwords (2.9\%) consist of phrases from publicly available sources.
-The exact percentage of passphrases used as passwords is not known, but this
-shows they are used by a lof of users while they have been studied relatively
-little.
 
@@
 To achieve this, we will use a probabilistic method which generates phrases in
-decreasing likelihood. An important factor in this design is the speed at which
-it can do this: most methods are not designed to iteratively list all
-possibilities in decreasing likelihood, which makes it difficult to do so using
-methods such as context-free grammar (as described by
-Weir\cite{probabilistic-thesis}, p. 84--85).
+decreasing likelihood. Which method to use, is part of the research. An
+important factor in this design is the speed at which it can do this: most
+methods are not designed to iteratively list all possibilities in decreasing
+likelihood, which makes it difficult to do so using methods such as
+context-free grammar (as described by Weir\cite{probabilistic-thesis}, p.
+84--85).
 
@@
 
+Possible probabilistic approaches are Markov chains, weighted context-free
+grammar and neural networks. These methods have been applied to password
+cracking in the past, and might be applied to passphrases as well. Markov
+chains is one example which has also been applied to
+passphrases\cite{sparell-simovits} with good results.
+
 \section{Scope}
@@
 such as birth dates of specific users. In our previous work\cite{own}, we found
-this to be not as effective as expected. It seems better to focus on the
-general method in this limited-time project.
+this to be not very effective in passphrases.
 
@@
 
-For software, we expect to use only free and open source products.
+For software, we expect to use only free and open source products. We will also
+publish our product as free and open source software.
 

### 2018-01-12 proposal/proposal.tex
Proposal: emphasize that only one prototype will be built

@@
 
-Labrande\cite{crackmeimfamous} attempted to crack hashes from the Korelogic
+Labrande\cite{crackmeimfamous} attempted to crack hashes from the KoreLogic
 dump (139 million entries) used a compilation of different public sources.
 Around 4.3 million passwords were cracked, of which 200 000 are 16 characters
-or longer. Public sources include quotes from WikiQuote and text written in
+or longer. Public sources include quotes from Wikiquote and text written in
 bold and italics on Wikipedia.
@@
 
-To achieve this, we will use a probabilistic method which generates phrases in
-decreasing likelihood. Which method to use, is part of the research. An
-important factor in this design is the speed at which it can do this: most
-methods are not designed to iteratively list all possibilities in decreasing
-likelihood, which makes it difficult to do so using methods such as
-context-free grammar (as described by Weir\cite{probabilistic-thesis}, p.
-84--85).
+To achieve this, we will use a probabilistic method. Which method to use is
+part of the research. An important factor in this design is the speed at which
+it can do this: most methods are not designed to find all possibilities in
+order of decreasing likelihood, which makes it difficult to do so using methods
+such as context-free grammar (as described by Weir\cite{probabilistic-thesis},
+p. 84--85).
 
@@
 	\item Which probabilistic methods can be used to generate output in decreasing order of likelihood?
-	\item Which probabilistic method produces the most likely passphrases?
-	\item How much data, both in terms of storage and random access memory, does this method require?
+	\item Which probabilistic methods produce realistic passphrases?
+	\item How much data, both in terms of storage and random access memory, does the chosen method require?
 \end{enumerate}
@@
 passphrases\cite{sparell-simovits} with good results.
+
+Due to the limited time, we will evaluate the possible approaches briefly and
+choose a method early on. The complete prototype will be built for only one of
+the possible methods.
 
@@
 
+\newpage
 \section{Ethical Considerations}

### 2018-01-29 final-product/combiner.py
Working version of n-gram passphrase generator (and related tools).

@@ -1,0 +1,89 @@
+#!/usr/bin/env python3
+
+import sys
+
+if '-h' in sys.argv or '--help' in sys.argv or len(sys.argv) == 1:
+    print(
+'''Usage:
+    combiner.py <n> <inf|min max> [-b <bonus>] <file ...>
+
+Combines n-gram files. Recommended to be run with pypy for speed.
+
+Arguments:
+  <n>
+    The n in n-gram, e.g. 3 for trigrams.
+
+  <inf|min max>
+    Either 'inf' to process all, or a min and max. Any words falling outside
+    the range will not be processed. This is useful to chop counting up into
+    pieces, such that it can fit into memory.
+
+  -b <bonus>
+    A bonus to be added (summed) to each entry in the first file. For example
+    if you have a really accurate source, such as previously cracked
+    passphrases, and a worse but more plentiful source such as n-grams parsed
+    from Wikipedia (there are many, but they're not very representative of
+    real passphrases), then you can apply the bonus to the passphrase n-grams
+    file to prioritize it.
+
+  <file ...>
+    One or more files. Must be in n-gram format (count + gram on each line).
+''')
+    exit(1)
+
+# Used to merge cracked passphrase ngrams and wikipedia based ngrams
+
+n = int(sys.argv[1])
+sys.argv = sys.argv[1:]
+
+bonus = 0
+if sys.argv[1] == '-b':
+    bonus = int(sys.argv[2])
+    sys.stderr.write('Applying bonus {} to first file\n'.format(bonus))
+    sys.argv = sys.argv[2:]
+
+if sys.argv[1] != 'inf':
+    processrange = (int(sys.argv[1]), int(sys.argv[2]))
+    sys.argv = sys.argv[2:]
+else:
+    processrange = (-float('inf'), float('inf'))
+    sys.argv = sys.argv[1:]
+
+total = {}
+
+overlap = 0
+notoverlap = 0
+skipped = 0
+
+for fileno, fname in enumerate(sys.argv[1:]):
+    sys.stderr.write('Processing file {}...\n'.format(fname))
+
+    for line in open(fname):
+        if line.strip() == '' or line[0] == '#' or ' ' not in line.strip():
+            continue
+
+        count, gram = line.strip().replace('!', '').replace('.', '').split(' ', 1)
+        count = int(count)
+        if len(gram.strip().replace('  ', ' ').split(' ')) != n:
+            skipped += 1
+            continue
+
+        if count >= processrange[0] and count <= processrange[1]:
[…]

### 2018-01-29 final-product/genphrases.py

@@ -1,0 +1,183 @@
+#!/usr/bin/env pypy
+
+import sys
+from lib import *
+
+if '-h' in sys.argv or '--help' in sys.argv or len(sys.argv) == 1:
+    print(
+'''Usage:
+
+  genphrases.py [-v] [-n <n>] [-m <minfreq>] <ngrams-file> <minlen> <maxlen>
+
+Generated phrases will be outputted to stdout. They are space-separated and
+meant to be used with the included hashcat rules.
+
+Arguments:
+  <ngrams-file>
+    A file with ngrams, generated by the companion tool, in sorted order. This
+    is loaded into memory (in memory, it is about 17.5x larger than on disk due
+    to both indexing and Python's overhead, unless you use -m).
+
+  <minlen>
+    Minimum number of words in each outputted phrase.
+
+  <maxlen>
+    Maximum number of words in each outputted phrase.
+
+  -m <minfreq>
+    Set a minimum frequency for the n-gram file, i.e. skip any lines with a
+    frequency of smaller than or equal to minfreq. Useful for limited memory.
+    A value of 2 often greatly reduces the memory footprint.
+
+  -v
+    Verbose mode. Messages are written to stderr.
+
+  -n <n>
+    Set the number of words in each n-gram to <n>, allowing the program to
+    treat lines with an incorrect number of words in the ngrams file as warning
+    instead of an error.
+
+Exit statuses:
+  0: OK
+  1: Error in arguments
+  2: Error in ngrams file
+''')
+    exit(1)
+
+minfreq = -1
+if '-m' in sys.argv:
+    i = sys.argv.index('-m')
+    minfreq = int(sys.argv[i + 1])
+    del sys.argv[i]
+    del sys.argv[i]  # not i+1 because it shifted
+
+verbose = False
+if '-v' in sys.argv:
+    verbose = True
+    del sys.argv[sys.argv.index('-v')]
+
+n = None
+''' The n value of the n-grams, e.g. 3 for trigrams. '''
+
+warnForIncorrectN = False
+if '-n' in sys.argv:
+    i = sys.argv.index('-n')
+    n = int(sys.argv[i + 1])
+    del sys.argv[i]
+    del sys.argv[i]  # not i+1 because it shifted
+    warnForIncorrectN = True
+
+index = {}
+''' Index of n-1 words, pointing to each previous n-gram starting with those words.
+For example:
+    index["there was"] = [1, 9] if
[…]

### 2018-01-29 final-product/lib.py

@@ -1,0 +1,5 @@
+import sys
+
+def err(msg):
+    sys.stderr.write(msg)
+

### 2018-01-30 paper/paper.tex
Paper WIP

\documentclass{article}

\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage[hyperref=true, natbib=true, style=numeric, backend=bibtex]{biblatex}
\usepackage{graphicx}
\bibliography{references.bib}

\setlength{\parindent}{0pt}
\setlength{\parskip}{\baselineskip}

\begin{document}

\title{Practical Passphrase Cracking}
\date{\today{}}
\author{Luc Gommans \\
	Radically Open Security
}

\maketitle

\section*{Abstract}

% TODO: write abstract

\section{Introduction}

% TODO: check section

Requirements for secure passwords are getting more and more stringent to stay
ahead of the increase in computational capacity of commonly accessible systems.
Using passphrases, users can approach the problem differently: form a string of
random, common words, instead of a complex sequence of characters. The most
famous example is `correct horse battery staple': four words with no coherence,
but many people in the field can recall this phrase without ever trying to
memorize it\cite{own}. Studies also show that users have fewer difficulties
when memorizing a passphrase than when memorizing equally-strong random
passwords\cite{behavioral-analysis}\cite{pwd-memorability}.

Password cracking tools are widely available, which help us understand the
strength of them. For passphrases, there is no commonly used approach using a
single tool or dataset, probably because there are very few of them.

In this work, we will attempt to create a program which generates likely
passphrases. It can be used in combination with traditional tools such as
Hashcat or John the Ripper. We will use a probabilistic method to create
realistic sentences and use rulesets of the traditional tool to do additional
manipulation such as uppercasing the first word, removing spaces, and other
operations that can be done using the optimized code of that tool.

In the next section, we will explore the state of the field. Section
\ref{sec:problemstatement}, Problem Statement, describes our expected
contribution. In order to obtain a set of passphrases to work with, we
reproduced Labrande's previous work in section \ref{sec:dataset}. Next, we
explore possible probabilistic methods in section \ref{sec:methods-comparison}.
Details about the implementation of the chosen method are set out in section
\ref{sec:implementation} and an analysis of its performance in section
\ref{sec:eval}. Finally, we describe our conclusions in section
\ref{sec:conclusions} and possible future work in section \ref{sec:futurework}.

\section{Related Work}

Sparell and Simovits\cite{sparell-simovits} created a passphrase cracking
program by using Markov chains on a word level. Markov chains are used in many
applications, such as next word prediction in keyboards on smartphones.
Similarly, one can use them to approximate the next most likely word in a
passphrase given a few starting words (which could be a dictionary walk). In
Sparell and Simovits' work, the LinkedIn hashdump of 2012 was used as one of
the test sets. This contains mostly passwords (not phrases), but is still
frequently used due to its large size of 64 million hashes. Of these, 21 000
phrases were cracked of between 10 and 20 characters in length.

In previous work, we\cite{own} generated passphrases from quotes of various
well-known persons and song lyrics. Using this method on the LinkedIn set, we
recovered 92 000 passphrases between 15 and 67 characters. A relatively large
amount of time was spent on collecting this data, which we think could be
optimized in future work. We also looked into whether users incorporate
personal data, such as lyrics specifically from a band they enjoy, but we have
not found this to be the common case.

Labrande\cite{crackmeimfamous} attempted to crack hashes from the Korelogic
dump (139 million entries) used a compilation of different public sources.
Around 4.3 million passwords were cracked, of which 200 000 are 16 characters
or longer. Public sources include quotes from WikiQuote and text written in
bold and italics on Wikipedia.

Bonneau and Shutova\cite{payphrase-properties} examined ,,patterns of human
choice in a passphrase-level authentication system deployed by Amazon''. They
,,tested the availability of [over] 100,000 possible phrases at Amazon's
registration page, which prohibits using any phrase already registered by
another user.'' This system allows passphrases as short as two words and
offered no explanation to users regarding what might be secure. Its interface
merely asks for a "phrase" (which is generally a string of {\it logical} words)
and even suggests two-word phrases to use. The conclusion drawn by this paper
is not surprising, given these circumstances: ,,users aren’t able to choose
phrases made of completely random words, but are influenced by the probability
of a phrase occurring in natural language. Examining the surprisingly weak
distribution of phrases in natural language, we can conclude that even 4-word
phrases probably provide less than 30 bits of security which is insufficient
against offline attack.'' We consider this low entropy estimation to be
specific for Amazon PayPhrase.

Weir\cite{weir} examined probabilistic methods for password cracking,
specifically Markov chains and probabilistic context-free grammar. Such
techniques might also be applied to passphrases. Besides examining different
methods, they also created an implementation using probabilistic context-free
grammar. This method cracked more passwords than John the Ripper (JtR), when
JtR is used in dictionary mode with the default word-mangling rules. It appears
that this method is well-suited for password cracking.

\section{Problem Statement}\label{sec:problemstatement}

Our main research question is: {\it How can software efficiently generate
likely passphrases, to be used in passphrase cracking?}

To achieve this, we will use a probabilistic method which generates phrases.
Which specific method to use, is part of this research. Probabilistic methods
can account for linguistic properties of phrases. It also allows users to
download a model, rather than a large list of (people's private) passphrases to
use directly. See section \ref{sec:methods-comparison} for a comparison of
possible methods.

Probabilistic methods require a dataset to be trained on. Because very few
systems use passphrases exclusively, there are no lists of commonly used
passphrases like there are for passwords. We will attempt to reproduce
Labrande's research\cite{crackmeimfamous} in order to create such a set. The
result of this attempt is described in section \ref{sec:dataset}.

We have also downloaded a large plaintext password dump, specifically the
"breach compilation" of 1.4 billion passwords, but this proved to mainly many
very insecure passwords and did not contain more than a few passphrases.

\section{Passphrase Dataset}\label{sec:dataset}

In order to train a probabilistic algorithm, a dataset is needed to train with.
Labrande\cite{crackmeimfamous} described cracking over 4 million entries from
the Korelogic\cite{korelogic} MD5 dataset, 200 000 of which are 16 or more
characters. This was done by downloading all articles of Wikipedia and
Wikiquote in various languages, plus a few other, smaller sources. Page titles,
list item entries, bold text, and italic text were used as phrases. Finally,
they used rules in John the Ripper to test variants of each phrase. They also
used rules which mangle the phrase into a password, for example by taking only
the first letter of each word in the phrase.

Because we are mainly interested in English passphrases, we downloaded only the
English Wikipedia and Wikiquote articles. For a set of hashes, we used the
LinkedIn as well as the Korelogic dump. This also provides a way of comparing
results with Sparell and Simovits'\cite{sparell-simovits} Markov chain-based
method, which they tested on the LinkedIn set. We processed the list in a
similar manner: extracting text in italics and bold, page titles, and list
items on Wikiquote. We used only a few phrase mangling rules, which are
described in section \ref{sec:implementation}. Since we are, at this point,
strictly interested in passphrases, we did not include many of the more
elaborate rules: rules such as taking the first letter of each word, or
appending random numbers, were not used. These defeat the purpose of a
passphrase, which is to be a set of common words with no punctuation or other
random tweaks to remember.

Labrande described the cracking process took a few weeks on "an outdated laptop
with a 2GHz processor", processing around 65 million unique entries in the
phrase dictionary. We extracted around 26 million unique entries, for a total
size of 690MiB. The difference in number of entries can be explained by the
fact that we did not download other languages. Processing these 38 million
entries, including the mangling rules, took around 10 minutes using Hashcat on
an Intel Core i7-3630QM.

From the LinkedIn set, this yielded 1.3 million results, of which 13 361 of 16
characters or longer. 26 345 cracked passwords include at least one space,
meaning they are probably passphrase-like even if they are shorter than sixteen
characters. In previous work\cite{own} we found that only about half the users
left spaces intact in passphrases, a number that is probably lower in a set
where users were not specifically queried about passphrases.

From the Korelogic MD5 set, we cracked 2.3 million hashes in 20 minutes, using
the same list of phrases and mangling rules. 146 742 were 16 characters or
longer and 311 815 included a space.

\section{Probabilistic Method Selection}\label{sec:methods-comparison}

Possible probabilistic approaches are Markov chains, weighted context-free
grammar and n-grams. Such methods have been applied to password cracking in
the past, and might be applied to passphrases as well. Markov chains is one
example which has also been applied to passphrases\cite{sparell-simovits} with
good results.

As far as we have been able to find, probabilistic context-free grammar has not
been applied to passphrase cracking. It proved to be a viable method in
Weir's\cite{weir} password cracking work. For passphrases, it could be adapted
to operate on whole words rather than individual characters. Based on texts and
previously cracked passphrases, rules could be derived which indicate for
example how likely it is for a noun to follow a verb.

In order to test whether this is a viable method, we attempted to classify the
type of word in cracked passphrases with a python program which used the
WordNet 3.0 database\cite{wordnet}. Because not all passphrases separate words
by spaces, the program would try to find words in phrases. The string
`correcthorsebattery' would result in the identified words `correct', `horse'
and `battery'. However, the WordNet database does not contain any conjugations
or plurals. In the string `hadadrink', it would not find the word `had' (it
only contains `to have'). The first known word would be `dad', after which it
would recognize `ink' as the next word. Incorrect identification of words
following an incorrect identification was extremely common. Particularly in
long words, it might recognize many small words incorrectly.

We evaluated the possibility of adding conjugations and plurals in some way, at
least for the most common words, but this quickly proved to be too large a task
to fit within the scope of this project.

While analysing the dataset, we saw that it contained many geographical names,
brands and person's names. Additionally, many of the phrases have the ring of a
catchphrase to them or form a common saying, for example "eye for an eye" or
"give me a break". While a probabilistic context-free grammar model might be
able to approximate those using large lists of names (and other words like
verbs, adjectives, etc.), n-grams capture parts of phrases and will inherently
become biased towards common phrases and constructions.

N-grams automatically include names and common constructions because they are
subsets of sentences, and implementations can be very fast because the
algorithm is fairly simple. The downside is that it has no memory: if it learns
`there are' and `are there' are common constructions, it will generate `are
there are', `there are there'. Markov chains have the same issue, but
context-free grammar does maintain more context (despite its name). We
chose to implement a proof of concept using n-grams because it is a novel
technique to generate passphrases and it seems fit for the purpose.

\section{Implementation}\label{sec:implementation}

,,[An] {\it n-gram} is a contiguous sequence of {\it n} items from a given
sample of text or speech.''\cite{wiki-ngram} In our case, each item is a word.
N-grams of the order two and three and also called {\it bigrams} and {\it
trigrams}, respectively.

We implemented a program in Python which splits a text on spaces and counts
each possible subsentence of {\it n} words. For example, in the sentence `lay
of the land', trigrams are `lay of the' and `of the land'. The number of times
an n-gram occurs is counted. Another Python program uses these n-grams to
generate phrases, using the most frequently occurring n-grams first. The source
code for these programs can be found in our git repository\cite{git}. The
output of the second program can be directly used in tools such as Hashcat.

The generator can make arbitrarily long phrases, as long as it can find a
possible continuation in the n-gram file (which is usually infinite because of
repetitions). The maximum length is therefore configurable. A logical value is
around eight to ten words: this is not long enough to output too many
repetitions while still generating fairly long phrases.

A bigrams file was initially generated from the passphrases we cracked in
section \ref{sec:dataset}. This yielded only 761 bigrams and 331 trigrams
because most passphrases are stripped of spaces. We attempted to use the
program which we used in section \ref{sec:methods-comparison} to find
concatenated words within phrases, splitting `batterystaple' into `battery' and
`staple'. Again we found that the identification errors were too great to be of
any use, even when using an additional dictionary\cite{wamerican} which
contains conjugations. (This additional dictionary was not used initially
because it does not have word type classifications.)

In order to supplement the system with more possible sentence continuations, we
used English Wikipedia articles. This yielded around 29 million bigrams and 47
million trigrams. We combined them with the n-grams from the cracked
passphrases, prioritizing the latter. This way, n-grams from Wikipedia are only
used supplementarily and n-grams from previously cracked passphrases are
tried first.

\section{Evaluation}\label{sec:eval}

\section{Conclusions}\label{sec:conclusions}

% TODO: compare Labrande's method with SparellSimovits' method (1.3M results of ours vs 20.5k of SS,
%   or 13k (ours) of 16+ chars vs 384 of SS.

\section{Future work}\label{sec:futurework}

\printbibliography

\end{document}

### 2018-02-01 paper/paper.tex
Paper: fixed references and changed title
Paper: wrote remaining sections; pushing binary pdf for convenience

@@
 
-\title{Practical Passphrase Cracking}
+\title{Probabilistic Passphrase Cracking}
 \date{\today{}}
@@
 
-% TODO: write abstract
+Passphrases are an alternative to passwords that are supposed to be easier to
+remember for users. For passwords, cracking tools are widely available, helping
+us understand their strengths and weaknesses. For passphrase cracking, there is
+currently no common approach which is universally seen as the most effective
+way.
+
+In this work, we attempted to find and implement the most effective method. We
+compared Markov chains, n-grams and a hybrid dictionary attack. The latter was
+found to be most effective, both in terms of efficiency and the number of
+results. Excluding the time required to build the dictionary from English
+Wikipedia and Wikiquote articles, it took only 10 minutes to crack over 2.3
+million unique passphrases by using a hybrid dictionary attack with hashcat.
+Future users could download such a dictionary without having to compile it
+manually from various sources.
+
+\newpage
 
@@
 
-% TODO: check section
-
 Requirements for secure passwords are getting more and more stringent to stay
 ahead of the increase in computational capacity of commonly accessible systems.
-Using passphrases, users can approach the problem differently: form a string of
-random, common words, instead of a complex sequence of characters. The most
+Using passphrases, users can approach the problem differently: instead of a
+complex sequence of characters, form a string of random, common words. The most
 famous example is `correct horse battery staple': four words with no coherence,
@@
 Password cracking tools are widely available, which help us understand the
-strength of them. For passphrases, there is no commonly used approach using a
-single tool or dataset, probably because there are very few of them.
-
-In this work, we will attempt to create a program which generates likely
-passphrases. It can be used in combination with traditional tools such as
-Hashcat or John the Ripper. We will use a probabilistic method to create
-realistic sentences and use rulesets of the traditional tool to do additional
-manipulation such as uppercasing the first word, removing spaces, and other
-operations that can be done using the optimized code of that tool.
+strength (or weakness) of passwords. For passphrases, there is no commonly used
+approach using a single tool or dataset, probably because there are very few of
+them.
 
@@
 passphrase given a few starting words (which could be a dictionary walk). In
-Sparell and Simovits' work, the LinkedIn hashdump of 2012 was used as one of
+Sparell and Simovits' work, the LinkedIn hash dump of 2012 was used as one of
 the test sets. This contains mostly passwords (not phrases), but is still
@@
 Labrande\cite{crackmeimfamous} attempted to crack hashes from the Korelogic
-dump (139 million entries) used a compilation of different public sources.
+dump (139 million MD5 entries) used a compilation of different public sources.
 Around 4.3 million passwords were cracked, of which 200 000 are 16 characters
-or longer. Public sources include quotes from WikiQuote and text written in
+or longer. Public sources include quotes from Wikiquote and text written in
 bold and italics on Wikipedia.
@@
 and even suggests two-word phrases to use. The conclusion drawn by this paper
-is not surprising, given these circumstances: ,,users aren’t able to choose
+is not surprising, given these circumstances: ,,users aren't able to choose
 phrases made of completely random words, but are influenced by the probability
@@
 methods, they also created an implementation using probabilistic context-free
-grammar. This method cracked more passwords than John the Ripper (JtR), when
-JtR is used in dictionary mode with the default word-mangling rules. It appears
-that this method is well-suited for password cracking.
+grammar. This method cracked more passwords than John the Ripper in dictionary
+mode with the default word-mangling rules. It appears that this method is
+well-suited for password cracking.
 
@@
 
-To achieve this, we will use a probabilistic method which generates phrases.
-Which specific method to use, is part of this research. Probabilistic methods
-can account for linguistic properties of phrases. It also allows users to
-download a model, rather than a large list of (people's private) passphrases to
-use directly. See section \ref{sec:methods-comparison} for a comparison of
-possible methods.
+To achieve this, we will use a probabilistic method which generates realistic
+phrases. Which specific method to use, is part of this research. We want to use
+a probabilistic method because they can account for linguistic properties of
+phrases. It also allows users to download a model, rather than a large list of
+phrases to use directly. In section \ref{sec:methods-comparison} we compare
+different methods.
+
+Our implementation will be used in combination with traditional password
+cracking tools such as Hashcat or John the Ripper, because those have fast
+implementations of hash functions and other helpful features. One such feature
+is rule sets, which we will use for generating variants of phrases to be tried.
+These rule languages can perform simple manipulations such as lowercasing each
+word, removing spaces, or appending an exclamation mark.
 
@@
 Labrande's research\cite{crackmeimfamous} in order to create such a set. The
-result of this attempt is described in section \ref{sec:dataset}.
-
-We have also downloaded a large plaintext password dump, specifically the
-"breach compilation" of 1.4 billion passwords, but this proved to mainly many
-very insecure passwords and did not contain more than a few passphrases.
+result of this attempt is described in the next section.
 
@@
 
-In order to train a probabilistic algorithm, a dataset is needed to train with.
-Labrande\cite{crackmeimfamous} described cracking over 4 million entries from
-the Korelogic\cite{korelogic} MD5 dataset, 200 000 of which are 16 or more
+In order to train a probabilistic algorithm, a dataset is needed to train on.
+Labrande described cracking over 4 million entries from the
+Korelogic\cite{korelogic} MD5 dataset, 200 000 of which are 16 or more
 characters. This was done by downloading all articles of Wikipedia and
@@
 
+We classify this attack as a hybrid dictionary attack, because of the
+similarity to traditional hybrid dictionary attacks on passwords. In suchs
+attacks, a large list of words is used together with mangling rules to generate
+possible permutations. In Labrande's variant, phrases are used instead of
+words, using a smart method of selecting likely phrases.
+
 Because we are mainly interested in English passphrases, we downloaded only the
@@
 results with Sparell and Simovits'\cite{sparell-simovits} Markov chain-based
-method, which they tested on the LinkedIn set. We processed the list in a
+method, which they tested on the LinkedIn set. We processed the data in a
 similar manner: extracting text in italics and bold, page titles, and list
@@
 described in section \ref{sec:implementation}. Since we are, at this point,
-strictly interested in passphrases, we did not include many of the more
-elaborate rules: rules such as taking the first letter of each word, or
-appending random numbers, were not used. These defeat the purpose of a
-passphrase, which is to be a set of common words with no punctuation or other
-random tweaks to remember.
+only interested in passphrases, we did not include many of the more elaborate
+rules: rules such as taking the first letter of each word or appending random
+numbers, were not used. Such random variations would defeat the purpose of a
+passphrase for the user, which is to be a set of common words with no
+punctuation or other random tweaks to remember.
 
-Labrande described the cracking process took a few weeks on "an outdated laptop
-with a 2GHz processor", processing around 65 million unique entries in the
-phrase dictionary. We extracted around 26 million unique entries, for a total
-size of 690MiB. The difference in number of entries can be explained by the
-fact that we did not download other languages. Processing these 38 million
-entries, including the mangling rules, took around 10 minutes using Hashcat on
-an Intel Core i7-3630QM.
+The results of this effort are described in section \ref{sub:dataset-results}
+(evaluation of the hybrid dictionary attack).
 
-From the LinkedIn set, this yielded 1.3 million results, of which 13 361 of 16
-characters or longer. 26 345 cracked passwords include at least one space,
-meaning they are probably passphrase-like even if they are shorter than sixteen
-characters. In previous work\cite{own} we found that only about half the users
-left spaces intact in passphrases, a number that is probably lower in a set
-where users were not specifically queried about passphrases.
-
-From the Korelogic MD5 set, we cracked 2.3 million hashes in 20 minutes, using
-the same list of phrases and mangling rules. 146 742 were 16 characters or
-longer and 311 815 included a space.
+Additionally, we have downloaded a large plain text password dump, specifically
+the `breach compilation' of 1.4 billion passwords. This proved to mainly
+consist of very insecure passwords and did not contain more than a few
+passphrases.
 
@@
 
-While analysing the dataset, we saw that it contained many geographical names,
+While analyzing the dataset, we saw that it contained many geographical names,
 brands and person's names. Additionally, many of the phrases have the ring of a
@@
 The generator can make arbitrarily long phrases, as long as it can find a
-possible continuation in the n-gram file (which is usually infinite because of
-repetitions). The maximum length is therefore configurable. A logical value is
-around eight to ten words: this is not long enough to output too many
-repetitions while still generating fairly long phrases.
+possible continuation in the n-gram file (which is typically infinite because
+of repetitions). The maximum length is therefore configurable. A logical value
+is around eight words: this does not output too many illogical repetitions
+while still generating fairly long phrases.
 
@@
 section \ref{sec:dataset}. This yielded only 761 bigrams and 331 trigrams
-because most passphrases are stripped of spaces. We attempted to use the
-program which we used in section \ref{sec:methods-comparison} to find
-concatenated words within phrases, splitting `batterystaple' into `battery' and
-`staple'. Again we found that the identification errors were too great to be of
-any use, even when using an additional dictionary\cite{wamerican} which
-contains conjugations. (This additional dictionary was not used initially
-because it does not have word type classifications.)
+because most users stripped their passphrases of spaces. We attempted to use
+the program which we used in section \ref{sec:methods-comparison} to find
+concatenated words within phrases, splitting e.g. `batterystaple' into
+`battery' and `staple'. Again we found that the identification errors were too
+great to be of any use, even when using an additional
+dictionary\cite{wamerican} which contains conjugations. (This dictionary was
+not used initially because it does not have word type classifications.)
 
@@
 
+Finally, we used mangling rules in Hashcat to create simple variants for each
+phrase. Hashcat can use multiple rule sets which are then combined, for example
+if in set 1 there are rules for both lowercasing and uppercasing a phrase and
+in set 2 there are rules for removing spaces, the resulting phrases for `A b'
+will be `a b', `A B', `ab' and `AB'. In our previous work, we learned what kind
+of alterations users make to their phrases. Based on that information, we
+created three sets of rules:
+
+\begin{enumerate}
+	\item \begin{itemize}
+		\item use as-is
+		\item lowercase all characters
+		\item capitalize the first letter, lowercase the rest
+		\item capitalize the first letter of each word, lowercase the rest (titlecase)
+	\end{itemize}
+
+	\item \begin{itemize}
+		\item use as-is
+		\item append a period
+		\item append an exclamation mark
+		\item append a question mark
+	\end{itemize}
+
+	\item \begin{itemize}
+		\item use as-is
+		\item remove all spaces
+	\end{itemize}
+\end{enumerate}
+
+These rules create 32 variants for each input.
+
+The system we used for all tests is a laptop with an Intel i7-3630QM CPU and
+8GiB RAM.
+
 \section{Evaluation}\label{sec:eval}
+
+In this section we will evaluate the performance of different passphrase
+cracking methods, as well as our n-gram implementation. We consider the number
+of entries they manage to crack as well as their efficiency.
+
+Because we reproduced Labrande's hybrid dictionary attack (with slight
+alterations) and cracked a set of hashes identical to those of Sparell and
+Simovits'\cite{sparell-simovits}, we can compare the performance of all methods
+directly.
+
+\subsection{Hybrid dictionary}\label{sub:dataset-results}
+
+Using the hybrid dictionary attack as described by Labrande (with slight
+alterations), we cracked 2.3 million hashes from the Korelogic MD5 set. 146 742
+were 16 characters or more. 311 815 included at least one space, meaning they
+are probably passphrase-like even if they are shorter than sixteen characters.
+
+This is only half as many as Labrande cracked: 4.2 million and 200 000 of 16
+characters or more (there are no statistics on the number of cracked entries
+with spaces). This difference could be explained by the fact that we were
+strictly after English passphrases and did not use a large number of mangling
+rules, and because we did not include other languages. Labrande used the
+English, German, French, Italian and Spanish versions of Wikipedia and
+Wikiquote.
+
+Note that in previous work\cite{own}, we found that only about half the users
+left spaces intact in passphrases. This number is probably even lower in a set
+where users were not specifically queried about passphrases. The number of
+passphrase-like entries probably more than twice as high as the number of
+entries we found with spaces.
+
+Of the LinkedIn set, we were able to crack 1.3 million entries, of which 13 361
+of 16 characters or more and 26 345 cracked passwords include a space.
+
+\subsubsection{Efficiency}
+
+Labrande's hybrid dictionary attack took ,,a few weeks on an outdated laptop
+with a 2GHz processor,'' using John the Ripper. When reproducing this attack,
+the process took around 10 minutes in Hashcat. Our dictionary was smaller (26
+million instead of 65 million entries) and used fewer mangling rules, which
+could explain the difference. Despite those limits, we still cracked half as
+many entries as Labrande. Despite the more than 2000-fold decrease in time, it
+still yielded good results.
+
+The storage requirements of this method are relatively large: our phrases
+file (with 26 million entries) is 690MiB. Since we had to generate this from
+Wikipedia, we also had to download all Wikipedia and Wikiquote articles, which
+are 26GiB as download and 135GiB uncompressed.
+
+Memory usage is not relevant for this attack, as each line can be read off of
+disk and is immediately passed onto Hashcat.
+
+\subsection{Markov chains}
+
+Sparell and Simovits cracked 25 000 entries from the LinkedIn set, of which 384
+of 16 characters or longer.
+
+\subsubsection{Efficiency}
+
+Sparell and Simovits described that generating phrases is the dominant part of
+the time consumption for their cracking process. They therefore wrote the
+output of their phrase generator to a file before inputting it into Hashcat,
+classifying files by different sources and passphrase lengths. Generation times
+differ from 16 minutes for a 2.6 million 10-character phrases file to 960 hours
+(40 days) for 8.5 billion 20-character phrases. The numbers in terms of phrases
+per second vary greatly, ranging from 22 500 to 2 459 phrases per second.
+
+The size of the source files, which form the Markov model, is not described;
+nor is the amount of RAM required for this method.
+
+\subsection{N-grams}
+
+% TODO: word choice (global): entries or records? Or something else?
+
+Using our passphrase generator with trigrams, we were able to crack 835 562
+entries of the Korelogic MD5 set, of which 32 911 are 16 characters or longer,
+and 70 198 contain a space. Of the LinkedIn set, we were able to crack 482 014
+entries, of which 3 782 are 16 characters or longer and 14 903 contain a space.
+
+Using bigrams, we were able to crack 342 145 entries of the Korelogic set, of
+which 9 148 are 16 characters or longer and 1 023 contain a space. Of the
+LinkedIn set, we were able to crack 190 117 entries, of which 1 468 are 16
+characters or longer and 1 449 contain a space.
+
+Both attacks produced a great number of short results, such as `thefault' or
+`wanderers'. After going through all phrases shorter than or equal to {\it n}
+words, the number of results dropped significantly.
+
+\subsubsection{Efficiency}
+
+The n-gram phrase generator first loads the data, which takes a few minutes and
+scales linearly with the n-gram file size. While running, it generates two to
+four million phrases per second on a single CPU core (the median is 3.3 million
+phrases per second). This saturated Hashcat's ability to process input: hashcat
+processes around ten million SHA-1 hashes per second, but applies a few rules.
+Since rules easily generate many more than three variants per phrase, the speed
+of our program is sufficient despite running on only a single core and being
+written in Python, an interpreted language.
+
+Memory usage for the generation process depends on the number of n-grams
+loaded: for 22 million trigrams (464MiB), it uses 5.3GiB of RAM. This is quite
+intensive, but required for fast lookups. Bigrams are much more efficient:
+because there are fewer combinations to be made, the input files are much
+smaller. The complete bigrams file contains 2.9 million entries (47MiB) and
+uses 800MiB RAM.
+
+A file with trigrams of 464MiB is not much smaller than our Wikipedia and
+Wikiquote phrase dictionary of 690MiB. Bigrams are more efficient, using 47MiB.
 
@@
 
-% TODO: compare Labrande's method with SparellSimovits' method (1.3M results of ours vs 20.5k of SS,
-%   or 13k (ours) of 16+ chars vs 384 of SS.
+We have explored and compared different methods for passphrase cracking,
+confirmed the results of Labrande's hybrid dictionary attack, and implemented a
+new method based on n-grams. Our conclusion is that the most effective attack
+at this time is the hybrid dictionary attack, even when considering the
+trade-off between storage requirements and practicality versus the number of
+results obtained. While the dictionary does need a relatively large amount of
+storage, particularly when generating or updating it, it is extremely efficient
+in terms of RAM and CPU.
+
+N-grams are less effective than expected. The most effective part of the
+cracking process were the phrases of {\it n} words or fewer, corresponding to
+the order of the n-gram. This means that literal hits were much more common
+than hits of combined n-grams, indirectly confirming again how powerful the
+dictionary attack is. It also indicates that passphrases are generally not
+normal, full sentences: despite training the model on both previously cracked
+passphrases and all English Wikipedia articles, sentences generated from those
+n-grams do not yield many results.
+
+Markov chains, as described by Sparell and Simovits, yielded the fewest results
+of the three methods we compared. Generating the passphrases was also slower
+than by using n-grams or a dictionary, probably because the algorithm for
+Markov chains is more complex.
+
+The source code for our software is available on Github\cite{git} under GPL
+version 3.0.
 
 \section{Future work}\label{sec:futurework}
+
+We have not been able to explore probabilistic context-free grammar fully due
+to the difficulties in classifying words. Weir was successful in this with
+passwords because he focused on a character level. For passphrases, which are
+composed of words instead of characters, this method is more complex. We
+recommend looking into other, possibly commercial options for a word
+classification database. We also do not recommend using previously cracked
+passphrases as training set: many are stripped of spaces, making it very
+difficult to programmatically split words for classification.
+
+Our n-gram attack can be improved in various ways. It might be more effective
+to store which n-grams were the start of a sentence. In our version, it would
+take an n-gram and append other, matching n-grams to complete the phrase.
+However, if the starting n-gram is not a logical way to start a sentence, all
+generated continuations are wasted. It would also be more memory- and
+storage-efficient to use a language with less memory overhead, and to identify
+words by tokens instead of indexing literal n-grams.
+
+Finally, we are curious how neural networks would perform for this purpose.
+Training a model is typically computationally intensive, but general purpose
+frameworks are available which make it easy to run such computations on a GPU
+and generate fairly large models on modest hardware.
 

### 2018-02-02 passphrase-cracker/genphrases.py
Genphrases: add benchmark mode, deduplicate shortphrases somewhat, fix large bug caused by reversed() (could apparently be iterated only once).

@@ -1,5 +1,9 @@
 #!/usr/bin/env pypy
 
+# Settings unconfigurable by command line arguments
+outhistsize = 5  # Higher is slower shortphrases, but fewer duplicates (and slightly more memory)
+
+
 import sys
 from lib import *
 
@@ -7,24 +11,24 @@ if '-h' in sys.argv or '--help' in sys.argv or len(sys.argv) == 1:
     print(
 '''Usage:
 
-  genphrases.py [-v] [-n <n>] [-m <minfreq>] <ngrams-file> <minlen> <maxlen>
+  genphrases.py [-v] [-b] [-n n] [-m minfreq] ngrams-file minlen maxlen
 
 Generated phrases will be outputted to stdout. They are space-separated and
 meant to be used with the included hashcat rules.
 
 Arguments:
-  <ngrams-file>
+  ngrams-file
     A file with ngrams, generated by the companion tool, in sorted order. This
     is loaded into memory (in memory, it is about 17.5x larger than on disk due
     to both indexing and Python's overhead, unless you use -m).
 
-  <minlen>
+  minlen
     Minimum number of words in each outputted phrase.
 
-  <maxlen>
+  maxlen
     Maximum number of words in each outputted phrase.
 
-  -m <minfreq>
+  -m minfreq
     Set a minimum frequency for the n-gram file, i.e. skip any lines with a
     frequency of smaller than or equal to minfreq. Useful for limited memory.
     A value of 2 often greatly reduces the memory footprint.
@@ -32,7 +36,10 @@ Arguments:
   -v
     Verbose mode. Messages are written to stderr.
 
-  -n <n>
+  -b
+    Benchmark mode. Only output the total number of generated phrases.
+
+  -n n
     Set the number of words in each n-gram to <n>, allowing the program to
     treat lines with an incorrect number of words in the ngrams file as warning
     instead of an error.
@@ -50,6 +57,11 @@ if '-m' in sys.argv:
     minfreq = int(sys.argv[i + 1])
     del sys.argv[i]
     del sys.argv[i]  # not i+1 because it shifted
+
+benchMode = False
+if '-b' in sys.argv:
+    benchMode = True
+    del sys.argv[sys.argv.index('-b')]
 
 verbose = False
 if '-v' in sys.argv:
@@ -90,6 +102,8 @@ if minlen > maxlen:
     exit(1)
 
[…]

### 2018-02-02 passphrase-cracker/combiner.py
Combiner.py: improved/fixed argument handling

@@ -1,4 +1,4 @@
-#!/usr/bin/env python3
+#!/usr/bin/env pypy
 
 import sys
 
@@ -7,7 +7,7 @@ if '-h' in sys.argv or '--help' in sys.argv or len(sys.argv) == 1:
 '''Usage:
     combiner.py <n> <inf|min max> [-b <bonus>] <file ...>
 
-Combines n-gram files. Recommended to be run with pypy for speed.
+Combines n-gram files. Outputs to stdout.
 
 Arguments:
   <n>
@@ -33,14 +33,16 @@ Arguments:
 
 # Used to merge cracked passphrase ngrams and wikipedia based ngrams
 
+bonus = 0
+if '-b' in sys.argv:
+    i = sys.argv.index('-b')
+    bonus = int(sys.argv[i + 1])
+    sys.stderr.write('Applying bonus {} to first file\n'.format(bonus))
+    del sys.argv[i]
+    del sys.argv[i]  # Not i+1 because it shifted
+
 n = int(sys.argv[1])
 sys.argv = sys.argv[1:]
-
-bonus = 0
-if sys.argv[1] == '-b':
-    bonus = int(sys.argv[2])
-    sys.stderr.write('Applying bonus {} to first file\n'.format(bonus))
-    sys.argv = sys.argv[2:]
 
 if sys.argv[1] != 'inf':
     processrange = (int(sys.argv[1]), int(sys.argv[2]))

### 2018-02-02 ngram-cracker/ngrams.py
Adding ngrams.py

@@ -1,0 +1,106 @@
+#!/usr/bin/env python3
+
+n = 2  # n-gram, n=?
+mincount = 3  # Only output grams that occurred >=mincount times
+maxmemusage = 1024 * 1024 * 5  # in KiB
+cleanupinterval = 5  # Check whether cleanup is needed every {cleanupinterval} seconds
+reducebelow = 0.85  # When over maxmemusage, reduce until maxmemusage*reducebelow
+
+import sys, os, time, gc
+
+if len(sys.argv) == 1 or '-h' in sys.argv or '--help' in sys.argv:
+    print('Usage: ngrams.py textfile | sort -n > output')
+    exit(1)
+
+def showOutput():
+    for gram in ngrams:
+        if ngrams[gram] >= mincount:
+            print('{} {}'.format(ngrams[gram], gram))
+
+
+def getmemusage():
+    # Memory usage in kB
+    f = open('/proc/{}/status'.format(os.getpid()))
+    memusage = int(f.read().split('VmRSS:')[1].split('\n')[0][:-3].strip())
+    f.close()
+    return memusage
+
+
+overusage = False  # Whether we're currently over our maxmemusage
+ngrams = {}
+ngramcount = 0
+linecount = 0
+gramsatlastcleanup = 0
+starttime = time.time()
+timesincecleanup = time.time()
+for line in open(sys.argv[1]):
+    words = line.strip().lower().split(' ')
+    if len(words) < n:
+        continue
+
+    for i in range(0, len(words) - (n - 1)):
+        nprime = n
+        gram = ''
+        space = ''
+
+        while nprime > 0:
+            """word = words[i + (n - nprime)]
+            if len(word) == 1 and word != 'i' and word != 'a' and not word.isdigit():
+                gram = ''
+                break"""
+
+            gram += space + words[i + (n - nprime)]
+            space = ' '
+            nprime -= 1
+
+        if gram not in ngrams:
+            ngrams[gram] = 1
+        else:
+            ngrams[gram] += 1
+
+        ngramcount += 1
+
+    linecount += 1
+
+    if not overusage and time.time() > timesincecleanup + cleanupinterval:
+        memusage = getmemusage()
+        speed = int(round(ngramcount / (time.time() - starttime) / 1000))
+        print('#N  Memory usage: {}MB  Avg speed: {}k grams/s  Total: {}k lines'.format(memusage / 1024, speed, linecount / 1000))
+
+        if memusage > maxmemusage:
[…]

### 2018-02-07 paper/paper.tex
Paper updated: added comparison table, and minor textual changes

@@
 \bibliography{references.bib}
-
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{\baselineskip}
 
@@
 
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{\baselineskip}
+
 \newpage
@@
 
-Password cracking tools are widely available, which help us understand the
-strength (or weakness) of passwords. For passphrases, there is no commonly used
+Password cracking tools, which help us understand the strength (or weakness) of
+passwords, are widely available. For passphrases, there is no commonly used
 approach using a single tool or dataset, probably because there are very few of
@@
 frequently used due to its large size of 64 million hashes. Of these, 21 000
-phrases were cracked of between 10 and 20 characters in length.
+phrases of between 10 and 20 characters in length were cracked.
 
@@
 Labrande\cite{crackmeimfamous} attempted to crack hashes from the Korelogic
-dump (139 million MD5 entries) used a compilation of different public sources.
+dump (139 million MD5 entries) using a compilation of different public sources.
 Around 4.3 million passwords were cracked, of which 200 000 are 16 characters
@@
 offered no explanation to users regarding what might be secure. Its interface
-merely asks for a "phrase" (which is generally a string of {\it logical} words)
+merely asks for a `phrase' (which is generally a string of {\it logical} words)
 and even suggests two-word phrases to use. The conclusion drawn by this paper
@@
 they used rules in John the Ripper to test variants of each phrase. They also
-used rules which mangle the phrase into a password, for example by taking only
+used rules that mangle the phrase into a password, for example by taking only
 the first letter of each word in the phrase.
@@
 
-Because we are mainly interested in English passphrases, we downloaded only the
+Since we are mainly interested in English passphrases, we downloaded only the
 English Wikipedia and Wikiquote articles. For a set of hashes, we used the
@@
 
-Additionally, we have downloaded a large plain text password dump, specifically
-the `breach compilation' of 1.4 billion passwords. This proved to mainly
-consist of very insecure passwords and did not contain more than a few
-passphrases.
+Additionally, a large plain text password dump was downloaded, specifically the
+`breach compilation' of 1.4 billion passwords. This proved to mainly consist of
+very insecure passwords and did not contain more than a few passphrases.
+
+Note that the downloaded dumps did not contain usernames, and any cracked hash
+values are kept confidential.
 
@@
 
-As far as we have been able to find, probabilistic context-free grammar has not
-been applied to passphrase cracking. It proved to be a viable method in
-Weir's\cite{weir} password cracking work. For passphrases, it could be adapted
-to operate on whole words rather than individual characters. Based on texts and
-previously cracked passphrases, rules could be derived which indicate for
-example how likely it is for a noun to follow a verb.
+So far, probabilistic context-free grammar has not been applied to passphrase
+cracking. It proved to be a viable method in Weir's\cite{weir} password
+cracking work. For passphrases, it could be adapted to operate on whole words
+rather than individual characters. Based on texts and previously cracked
+passphrases, rules could be derived which indicate for example how likely it is
+for a noun to follow a verb.
 
@@
 
-While analyzing the dataset, we saw that it contained many geographical names,
-brands and person's names. Additionally, many of the phrases have the ring of a
-catchphrase to them or form a common saying, for example "eye for an eye" or
-"give me a break". While a probabilistic context-free grammar model might be
-able to approximate those using large lists of names (and other words like
-verbs, adjectives, etc.), n-grams capture parts of phrases and will inherently
-become biased towards common phrases and constructions.
+While analyzing the dataset, it was found that it contained many geographical
+names, brands and people's names. Additionally, many of the phrases have the
+ring of a catchphrase to them or form a common saying, for example "eye for an
+eye" or "give me a break". While a probabilistic context-free grammar model
+might be able to approximate those using large lists of names (and other words
+like verbs, adjectives, etc.), n-grams capture parts of phrases and will
+inherently become biased towards common phrases and constructions.
 
 N-grams automatically include names and common constructions because they are
-subsets of sentences, and implementations can be very fast because the
-algorithm is fairly simple. The downside is that it has no memory: if it learns
-`there are' and `are there' are common constructions, it will generate `are
-there are', `there are there'. Markov chains have the same issue, but
-context-free grammar does maintain more context (despite its name). We
-chose to implement a proof of concept using n-grams because it is a novel
-technique to generate passphrases and it seems fit for the purpose.
+subsets of sentences, and implementations can be very fast due to the
+simplicity of the algorithm. The downside is that it has no memory: if it
+learns `there are' and `are there' are common constructions, it will generate
+`are there are', `there are there'. Markov chains have the same issue, but
+context-free grammar does maintain more context (despite its name). We chose to
+implement a proof of concept using n-grams because it is a novel technique to
+generate passphrases and it seems fit for the purpose.
 
-\section{Implementation}\label{sec:implementation}
+\section{N-gram implementation}\label{sec:implementation}
 
@@
 
-We implemented a program in Python which splits a text on spaces and counts
+A program was implemented in Python which splits a text on spaces and counts
 each possible subsentence of {\it n} words. For example, in the sentence `lay
@@
 
-A bigrams file was initially generated from the passphrases we cracked in
-section \ref{sec:dataset}. This yielded only 761 bigrams and 331 trigrams
+A bigrams file was initially generated from the passphrases that were cracked
+in section \ref{sec:dataset}. This yielded only 761 bigrams and 331 trigrams
 because most users stripped their passphrases of spaces. We attempted to use
-the program which we used in section \ref{sec:methods-comparison} to find
+the program mentioned in section \ref{sec:methods-comparison} to find
 concatenated words within phrases, splitting e.g. `batterystaple' into
@@
 used English Wikipedia articles. This yielded around 29 million bigrams and 47
-million trigrams. We combined them with the n-grams from the cracked
+million trigrams. Those were combined with the n-grams from the cracked
 passphrases, prioritizing the latter. This way, n-grams from Wikipedia are only
-used supplementarily and n-grams from previously cracked passphrases are
-tried first.
+used supplementarily and n-grams from previously cracked passphrases are tried
+first.
 
-Finally, we used mangling rules in Hashcat to create simple variants for each
-phrase. Hashcat can use multiple rule sets which are then combined, for example
-if in set 1 there are rules for both lowercasing and uppercasing a phrase and
-in set 2 there are rules for removing spaces, the resulting phrases for `A b'
-will be `a b', `A B', `ab' and `AB'. In our previous work, we learned what kind
-of alterations users make to their phrases. Based on that information, we
-created three sets of rules:
+Finally, mangling rules were used in Hashcat to create simple variants for each
+phrase. Hashcat can use multiple rule sets which are then combined. If, for
+example, if in set 1 there are rules for both lowercasing and uppercasing a
+phrase and in set 2 there are rules for removing spaces, the resulting phrases
+for `A b' will be `a b', `A B', `ab' and `AB'. In our previous work, we learned
+what kind of alterations users make to their phrases. Based on that information,
+we created three sets of rules:
 
@@
 
-The system we used for all tests is a laptop with an Intel i7-3630QM CPU and
-8GiB RAM.
+The system used for all tests is a laptop with an Intel i7-3630QM CPU and 8GiB
+RAM.
 
@@
 
-In this section we will evaluate the performance of different passphrase
-cracking methods, as well as our n-gram implementation. We consider the number
-of entries they manage to crack as well as their efficiency.
+In this section the performance of different passphrase cracking methods is
+evaluated, considering both the number of entries they manage to crack and
+their efficiency.
 
@@
 alterations) and cracked a set of hashes identical to those of Sparell and
-Simovits'\cite{sparell-simovits}, we can compare the performance of all methods
-directly.
+Simovits'\cite{sparell-simovits}, the performance of all methods can be
+compared directly.
 
@@
 Of the LinkedIn set, we were able to crack 1.3 million entries, of which 13 361
-of 16 characters or more and 26 345 cracked passwords include a space.
+are 16 characters or more and 26 345 cracked passwords include a space.
 
@@
 
-The storage requirements of this method are relatively large: our phrases
-file (with 26 million entries) is 690MiB. Since we had to generate this from
-Wikipedia, we also had to download all Wikipedia and Wikiquote articles, which
-are 26GiB as download and 135GiB uncompressed.
+The storage requirements of this method are relatively large: our phrases file
+(with 26 million entries) is 690MiB. Since this was generated from Wikipedia,
+all Wikipedia and Wikiquote articles also had to be downloaded, which total
+26GiB as download and 135GiB uncompressed.
 
@@
 Sparell and Simovits cracked 25 000 entries from the LinkedIn set, of which 384
-of 16 characters or longer.
+are 16 characters or longer.
 
@@
 
-Using our passphrase generator with trigrams, we were able to crack 835 562
+Using the passphrase generator with trigrams, it was able to crack 835 562
 entries of the Korelogic MD5 set, of which 32 911 are 16 characters or longer,
-and 70 198 contain a space. Of the LinkedIn set, we were able to crack 482 014
+and 70 198 contain a space. Of the LinkedIn set, it was able to crack 482 014
 entries, of which 3 782 are 16 characters or longer and 14 903 contain a space.
 
-Using bigrams, we were able to crack 342 145 entries of the Korelogic set, of
+Using bigrams, it was able to crack 342 145 entries of the Korelogic set, of
 which 9 148 are 16 characters or longer and 1 023 contain a space. Of the
-LinkedIn set, we were able to crack 190 117 entries, of which 1 468 are 16
+LinkedIn set, it was able to crack 190 117 entries, of which 1 468 are 16
 characters or longer and 1 449 contain a space.
@@
 Wikiquote phrase dictionary of 690MiB. Bigrams are more efficient, using 47MiB.
+
+\subsection{Overview}
+
+Table \ref{tab:comparison} gives an overview of the performance of the
+different methods. The column headers `Dict-1' and `Dict-2' indicate the hybrid
+dictionary attack, Labrande's and ours, respectively. The rows labeled `16+'
+indicate the number of results which are 16 characters or longer.
+
+Note that for our hybrid dictionary attack, the number of generated phrases per
+second is estimated at 10 million. Contrarily to all other methods, this method
+is IO-bound instead of CPU-bound. The amount of RAM required is considered
+negligible, as it only requires IO buffers.
+
+\begin{table}[H]
+	\begin{tabular}{| l | r | r | r | r | }
+	\hline                & {\bf Dict-1} & {\bf Dict-2} & {\bf Markov} & {\bf 3-grams} \\
+	\hline LinkedIn       & n/a          &  1 300 000   &    25 000    &   482 014     \\
+	\hline LinkedIn 16+   & n/a          &     13 361   &       384    &     3 782     \\
+	\hline Korelogic      & 4 200 000    &  2 300 000   & n/a          &   835 562     \\
+	\hline Korelogic 16+  &   200 000    &    146 742   & n/a          &    32 911     \\
+	\hline Phrases/second & n/a          & 10 000 000   &    22 500    & 3 300 000     \\
+	\hline RAM            & n/a          & negligible   & n/a          & 5.3GiB        \\
+	\hline Storage        & n/a          &     690MiB   & n/a          & 464MiB        \\
+	\hline
+	\end{tabular}
+	\caption{Performance overview}
+	\label{tab:comparison}
+\end{table}
 
@@
 passwords because he focused on a character level. For passphrases, which are
-composed of words instead of characters, this method is more complex. We
-recommend looking into other, possibly commercial options for a word
-classification database. We also do not recommend using previously cracked
-passphrases as training set: many are stripped of spaces, making it very
-difficult to programmatically split words for classification.
+composed of words instead of characters, this method is more complex. It is
+recommended to look into other, possibly commercial options for a word
+classification database. It is also recommended to avoid using previously
+cracked passphrases as training set: many are stripped of spaces, making it
+very difficult to programmatically split words for classification.
 
@@
 
+\newpage
 \printbibliography

### 2018-02-07 paper/paper.tex
Paper: committing new pdf, and added a comma in the text

@@
 distribution of phrases in natural language, we can conclude that even 4-word
-phrases probably provide less than 30 bits of security which is insufficient
+phrases probably provide less than 30 bits of security, which is insufficient
 against offline attack.'' We consider this low entropy estimation to be

### 2018-02-07 paper/paper.tex
Paper: spell-checked (committing binary pdf and tex), resolved a todo

@@
 Wikipedia and Wikiquote articles, it took only 10 minutes to crack over 2.3
-million unique passphrases by using a hybrid dictionary attack with hashcat.
+million unique passphrases by using a hybrid dictionary attack with Hashcat.
 Future users could download such a dictionary without having to compile it
@@
 We classify this attack as a hybrid dictionary attack, because of the
-similarity to traditional hybrid dictionary attacks on passwords. In suchs
+similarity to traditional hybrid dictionary attacks on passwords. In such
 attacks, a large list of words is used together with mangling rules to generate
@@
 
-% TODO: word choice (global): entries or records? Or something else?
-
 Using the passphrase generator with trigrams, it was able to crack 835 562
@@
 four million phrases per second on a single CPU core (the median is 3.3 million
-phrases per second). This saturated Hashcat's ability to process input: hashcat
+phrases per second). This saturated Hashcat's ability to process input: Hashcat
 processes around ten million SHA-1 hashes per second, but applies a few rules.
